---
title: "Sentiment Analysis Project"
author: "Noblezada, Camasa, Cabia"
date: "2024-12-13"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library(tidyverse)
library(tidytext)
library(dplyr)
library(stringr)
library(ggplot2)
library(sentimentr)
library(lubridate)
```

```{r}
# Load the dataset
tweetsDF <- read_csv("/cloud/project/ProjectDS/tweetsDF.csv")
```
# Cleaning the tweet text
```{r}

tweetsDF <- tweetsDF %>%
  mutate(
    text = text %>%
      iconv(from = "UTF-8", to = "ASCII//TRANSLIT", sub = "") %>% # Remove non-ASCII characters
      tolower() %>% # Convert to lowercase
      str_remove_all("https\\S+") %>% # Remove URLs
      str_remove_all("[#\\n]") %>% # Remove hashtags and newlines
      str_remove_all("[@?]\\S+") %>% # Remove mentions
      str_remove_all("\\?") %>% # Remove question marks
      str_remove_all("\\b\\d{2}\\.\\d{2}\\.\\d{4}\\b") %>% # Remove dates in dd.mm.yyyy format
      str_remove_all("<a href=httptwitter.comdownloadiphone rel=nofollow>twitter for iphone<a>") %>% # Remove specific substrings
      str_remove_all("<a href=([^>]*?) rel=nofollow>([^<]*?)<a>") %>%
      str_remove_all("<a href=httptwitter.comdownloadandroid rel=nofollow>twitter for android<a>") %>%
      str_remove_all("<a href= rel=nofollow>twitter web app<a>") %>%
      str_remove_all("30102022") %>% # Remove specific date
      str_squish() # Remove extra whitespace
  )
tweetsDF <- tweetsDF %>%
  mutate(date = ymd_hms(created)) %>%  
  mutate(hour = hour(date)) 

print(tweetsDF)

TweetsDF <- head(tweetsDF, 1000)

write.csv(TweetsDF, "CleanTweets.csv", row.names = FALSE)
```

```{r}
colnames(tweetsDF)
```

# Trend Analysis
```{r}

# Load and preprocess the dataset
tweets_df <- read.csv("/cloud/project/ProjectDS/tweetsDF.csv")
tweets_df$created <- ymd_hms(tweets_df$created)

# Group tweets by hour and count
tweets_per_time <- tweets_df %>%
  mutate(hour = floor_date(created, "hour")) %>%
  count(hour)

# Plot the trend analysis
ggplot(tweets_per_time, aes(x = hour, y = n)) +
  geom_line(color = "red") +
  labs(
    title = "Tweet Activity Over Time",
    x = "Time",
    y = "Number of Tweets"
  ) +
  theme_minimal()

```
# Observations:
# - Tweets were minimal before midnight on October 29.
# - A significant spike occurred after 10:30 PM on October 29, coinciding with the Itaewon tragedy.
# - Activity peaked in the early morning of October 30, declined through the morning, and rose again in the evening.


# SENTIMENT ANALYSIS
```{r}

# Perform sentiment analysis on tweets to examine emotional responses.

# Load sentiment lexicon and preprocess tweets
sentiments <- get_sentiments("bing")
tweet_tokens <- tweets_df %>%
  unnest_tokens(word, text) %>%
  inner_join(sentiments, by = "word")

# Count sentiments and calculate percentages
sentiment_counts <- tweet_tokens %>%
  count(sentiment) %>%
  mutate(percentage = n / sum(n) * 100)

# Print sentiment counts
print(sentiment_counts)

# Plot sentiment analysis
ggplot(sentiment_counts, aes(x = sentiment, y = percentage, fill = sentiment)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Sentiment Analysis of Tweets",
    x = "Sentiment",
    y = "Percentage"
  ) +
  theme_minimal()


```
# Observations:
# - The plot shows the proportion of positive and negative sentiments in the tweets.
# - Helps to understand the overall emotional tone of the conversations.

```{r}
# Analyze the dataset to determine the distribution of tweets across different source platforms.
# Group tweets by source and calculate the count for each source
tweet_source_analysis <- tweets_df %>%
  group_by(tweetSource) %>%
  summarise(Number_of_Tweets = n()) %>%
  arrange(desc(Number_of_Tweets))

# Plot the number of tweets by source
ggplot(tweet_source_analysis, aes(x = reorder(tweetSource, -Number_of_Tweets), y = Number_of_Tweets)) +
  geom_bar(stat = "identity", fill = "green") +
  labs(
    title = "Tweet Volume by Source Platform",
    x = "Source Platform",
    y = "Number of Tweets"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```
# Observations:
# - The bar graph illustrates the distribution of tweets by source platform.
# - The x-axis displays various platforms (e.g., iPhone, Android, others), while the y-axis represents the total tweet count from each source.
# - iPhone and Android dominate as the most frequently used platforms for engaging with the content.


